---
title: 'Assignment 2 - EDS 241: Environmental Policy Evaluation'
author: "Juliet Cohen"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output:
  pdf_document:
    toc: false
    number_sections: yes
editor_options: 
  markdown: 
    wrap: 72
---

The goal of this assignment is to provide a simple test of whether the
effects of air quality regulations are the same across locations with
different racial mix. To this end you will test if the NOx Budget
Program, a cap-and-trade market for nitrogen oxides (NOx) emissions from
power plants lead to similar effects in counties that are predominantly
white versus counties that are predominantly African American. The data
are a subset of data sample I used in the following paper:
<https://olivierdeschenes.weebly.com/uploads/1/3/5/0/135068654/defensive-investmentsand-the-demans-for-air-quality.pdf>.
You can also get more information on the NOx Budget Program, here:
<https://www.epa.gov/airmarkets/nox-budget-trading-program>

The data included in the file NBP.xls, which is available on
Gauchospace, are: fips (fips code identifying each county), NBP
(indicator =1 if the county was regulated under the NOx Budget Program),
`PctBlack` (fraction of the county population that is African American),
and Dnox_masstons (change in annual NOx emissions from all power plants
in a county between 2000 and 2008 (in tons)). Note that the NBP market
was in effect in 212 of the 485 counties in the sample from 2003 to
2008, so the 2008-2000 change give us a sense of the program's effect on
emissions. If emissions of NOx from power plants declined in a county,
then Dnox_masstons should be negative.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

### Load Packages

```{r}
library(here)
library(dplyr)
library(tidyverse)
library(estimatr)
library(modelsummary)
library(knitr)
library(readxl)
```

### Read in Data

```{r}
data <- readxl::read_excel(here("NBP.xls"))

# check for NA values
#map(data, ~sum(is.na(.)))
# there are none
```

## (a) Make a histogram depicting the distribution of Dnox_masstons.

```{r}
emissions_histogram <- ggplot(data = data, aes(x = Dnox_masstons)) +
  geom_histogram(bins = 50, color = "black", fill = "aquamarine3") +
  ggtitle("Change in Annual Power Plant NOx Emissions (2000-2008)") +
   xlab("NOx Emissions (tons)") + 
   ylab("Count") +
   theme(panel.background = element_blank(),
         axis.title.x = element_text(color = "black", size = 13),
         axis.text.x = element_text(face = "bold", color = "black", size = 15),
         axis.title.y = element_text(color = "black", size = 15),
         axis.text.y = element_text(face = "bold", color = "black", size = 12),
         plot.title = element_text(color="black", size = 15, face = "bold"),
         panel.border = element_rect(colour = "black", fill = NA, size = 2)) +
  scale_y_continuous(breaks = seq(0, 160, by = 20)) +
  scale_x_continuous(breaks = seq(-60, 6, by = 10))

emissions_histogram
```

## (b) Create an indicator =1 if the county has `PctBlack` above the sample median, and =0 otherwise (in the rest of the assignment, I refer to this variable as 'D'). What is the average of `PctBlack` for counties above the median, i.e. counties for which D=1?

```{r}
# find median of the PctBlack values
median_pct_blk <- median(data$PctBlack)
median_pct_blk
```

```{r}
data_blk_med <- data %>% 
  mutate(abv_bel_med_blk = case_when(
    PctBlack > median_pct_blk ~ 1,
    PctBlack <= median_pct_blk ~ 0))

data_blk_med_summary <- data_blk_med %>% 
  group_by(abv_bel_med_blk) %>% 
  summarise(mean = mean(PctBlack))

data_blk_med_summary$mean[1]
data_blk_med_summary$mean[2]
```

\noindent The average of `PctBlack` for counties above the median (D=1)
is `r data_blk_med$mean[2]`.

## (c) Estimate a regression of Dnox_masstons on NBP. Interpret the estimated intercept and the coefficient on NBP.

```{r}
nox_nbp_model <- lm_robust(formula = Dnox_masstons ~ NBP, data = data)

#nox_nbp_model$coefficients[1]
#nox_nbp_model$coefficients[2]

nox_nbp_model_table <- tidy(nox_nbp_model)
nox_nbp_model_table %>%
  select(term, estimate, std.error, p.value, conf.low, conf.high) %>% 
  kable()
nox_nbp_model_table
```

\noindent - For the regression of `Dnox_masstons` on `NBP`, the
estimated intercept is `r nox_nbp_model$coefficients[1]`. This means
that for the average county without these NBP Budget Program regulations (NBP is held constant at 0), the amount of change in annual NOx
emissions from all power plants between 2000 and 2008 was
`r nox_nbp_model$coefficients[1]` tons. Since this number is negative,
we can conclude that annual NOx emissions decreased by that amount when
all else is held constant.

\noindent - The estimated coefficient on NBP is
`r nox_nbp_model$coefficients[2]`. This means that for the average county with these regulations
under the NOx Budget Program, the amount of change in annual
NOx emissions from all power plants between 2000 and 2008
was `r nox_nbp_model$coefficients[2]` tons. Since this number is
negative, we can conclude that the NOx emissions decreased by that
amount when counties were regulated but all else is held constant.

```{r, include = FALSE, eval = FALSE}
ggplot(data, aes(y = Dnox_masstons, x = NBP)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)
```

## (d) Create an interaction between the variables NBP and D. Estimate a regression of Dnox_masstons on NBP, D, and this interaction. Interpret each estimated regression coefficient, including the intercept.

```{r, include = FALSE, eval = FALSE}
#nox_nbp_d_model$coefficients[1]
#nox_nbp_d_model$coefficients[2]
#nox_nbp_d_model$coefficients[3]
#nox_nbp_d_model$coefficients[4]
```

```{r}
nox_nbp_d_model <- lm_robust(formula = Dnox_masstons ~ NBP + abv_bel_med_blk + NBP:abv_bel_med_blk, data = data_blk_med)

nox_nbp_d_model_table <- tidy(nox_nbp_d_model)
nox_nbp_d_model_table %>%
  select(term, estimate, std.error, p.value, conf.low, conf.high) %>% 
  kable()
```

\noindent - The intercept is `r nox_nbp_d_model$coefficients[1]`, which
represents the average county's change in annual NOx emissions from all power plants between 2000 and 2008 (in tons) when the county was not regulated
under the NOx Budget Program and the county's `PctBlack` is below the
sample median, meaning the fraction of the county population that is
African American is below the sample median.

\noindent - The estimated regression coefficient for NBP is
`r nox_nbp_d_model$coefficients[2]`, which represents the average county's change in
annual NOx emissions from all power plants between 2000 and
2008 (in tons) when the county is regulated under the NOx Budget Program
and the county's `PctBlack` is below the sample median, meaning the
fraction of the county population that is African American is below the
sample median.

\noindent - The estimated regression coefficient for D (which is
represented by the variable `abv_bel_med_blk` in my code) is
`r nox_nbp_d_model$coefficients[3]`, which represents the average county's change in
annual NOx emissions from all power plants between 2000 and
2008 (in tons) when the county is not regulated under the NOx Budget
Program and the county's `PctBlack` is above the sample median, meaning
the fraction of the county population that is African American is above
the sample median.

\noindent - The estimated regression coefficient for the interaction
between NBP and D is `r nox_nbp_d_model$coefficients[4]`,
which represents the difference in the average county's change in annual NOx emissions
from all power plants between 2000 and 2008 (in tons) when
counties are regulated under the NOx Budget Program and the
`PctBlack` is above the sample median, meaning the fraction of the county
population that is African American is above the sample median, compared
to the average county's change in annual NOx emissions from all power plants between 2000 and 2008 (in tons) when there are regulations under
the NOx Budget Program and the county's `PctBlack` is below the sample
median, meaning the fraction of the county population that is African
American is below the sample median. This interaction term is the difference in the effect of regulation between predominantly black and predominantly white counties. We are investigating if the regulations bring the same benefits the these two categories of counties from an environmental justice perspective. Because this interaction term is not 0, we know there is indeed a difference for predominantly black and white communities, and because this interaction term is positive, we know that predominantly black communities are not getting as much benefit from these regulations compared to white communities. Emissions are improving for both predominantly white and predominantly black counties, but not as much for predominantly black counties.


```{r, include = FALSE, eval = FALSE}
# See the regression specification slides for an example of each beta's meaning.

ggplot(data_blk_med, aes(y = Dnox_masstons, x = NBP, color = abv_bel_med_blk)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)
```

## (e) What is the predicted Dnox_masstons in a county that was not regulated under NBP and where PctBlack is above the sample median (i.e., where D=1)? Report the 95% confidence interval for this prediction. Make sure to use "heteroskedasticity-robust" standard errors.

```{r}
not_reg_abv_blk <- data.frame(NBP = c(0), abv_bel_med_blk = c(1))

CI <- predict(object = nox_nbp_d_model, newdata = not_reg_abv_blk, se.fit = TRUE, interval = "confidence")
CI
```

\noindent The predicted Dnox_masstons in a county that is not regulated
under NBP and where `PctBlack` is above the sample median is -5.006106 tons,
which represents the change in annual NOx emissions from all power
plants in a county between 2000 and 2008 (in tons). The 95% confidence
interval for this prediction is [-6.440065, -3.572147].

